## From a partitioned parquet file

Let's now look at how to perform queries on **partitioned files**.   

The structure of partitioned files on the disk is as follows:

```{r}
#| label: viz-dir-tree

fs::dir_tree(path = "Datasets/DataMultiTypes/")
```

For this comparison, we will use :

- For **arrow (lazy)**, the `arrow::open_dataset()` method
- For **dplyr (duckdb)**, the ` DBI::dbConnect`, `dplyr::tbl()` and `arrow::read_parquet()` methods
- For **polars (lazy)**, the `pl$scan_parquet()` method


::: {.panel-tabset}

## arrow (lazy)

```{r}
#| label: partitioned-parquet-arrow-lazy-benchmarking

partitioned_parquet_arrow_lazy <- function(variables) {
  
  result <- arrow::open_dataset(
    "Datasets/DataMultiTypes/",
    partitioning = arrow::schema(colFactor = arrow::utf8())) |>
    
    mutate(
      # Conversion of 2 columns to Date format
      colDate1 = as.Date(colDate1),
      colDate2 = as.Date(colDate2)
    ) |>
    # Filter rows
    filter(
      colInt>2000 & colInt<8000
    ) |>
    # Grouping and aggregation
    group_by(colString) |> 
    summarise(
      min_colInt = min(colInt),
      mean_colInt = mean(colInt),
      mas_colInt = max(colInt),
      min_colNum = min(colNum),
      mean_colNum = mean(colNum),
      max_colNum = max(colNum)
    ) |> 
    collect()
  
  return(result)
  
}
tic()
partitioned_parquet_arrow_lazy() 
toc()
```

## dplyr (duckdb)

```{r}
#| label: partitioned-parquet-duckdb-benchmarking

# library(dbplyr)

partitioned_parquet_dplyr_duckdb <- function(variables) {
  
  con <- DBI::dbConnect(duckdb::duckdb())
  
  result <- tbl(con, "read_parquet('Datasets/DataMultiTypes/*/*.parquet', hive_partitioning=1)") |>
    
    mutate(
      # Conversion of 2 columns to Date format
      colDate1 = as.Date(colDate1),
      colDate2 = as.Date(colDate2)
    ) |>
    # Filter rows
    filter(
      colInt>2000 & colInt<8000
    ) |>
    # Grouping and aggregation
    group_by(colString) |> 
    summarise(
      min_colInt = min(colInt, na.rm = TRUE),
      mean_colInt = mean(colInt, na.rm = TRUE),
      mas_colInt = max(colInt, na.rm = TRUE),
      min_colNum = min(colNum, na.rm = TRUE),
      mean_colNum = mean(colNum, na.rm = TRUE),
      max_colNum = max(colNum, na.rm = TRUE)
    ) |> 
    collect()
  
  DBI::dbDisconnect(con)
  return(result)
}
tic()
partitioned_parquet_dplyr_duckdb() 
toc()
```

## polars (lazy)

```{r}
#| label: partitioned-parquet-polars-lazy-benchmarking

partitioned_parquet_polars_lazy <- function(variables) {
  
  result <- pl$scan_parquet(file = "Datasets/DataMultiTypes.parquet")$
    # Conversion of 2 columns to Date format
    with_columns(
      pl$col("colDate1")$str$strptime(pl$Date, "%F %T", strict = FALSE),
      pl$col("colDate2")$str$strptime(pl$Date, "%F %T", strict = FALSE)
    )$
    # Filter rows
    filter(
      pl$col("colInt")>2000 & pl$col("colInt")<8000
    )$
    # Grouping and aggregation
    groupby(
      "colString")$
    agg(
      pl$col("colInt")$min()$alias("min_colInt"),
      pl$col("colInt")$mean()$alias("mean_colInt"),
      pl$col("colInt")$max()$alias("max_colInt"),
      pl$col("colNum")$min()$alias("min_colNum"),
      pl$col("colNum")$mean()$alias("mean_colNum"),
      pl$col("colNum")$max()$alias("max_colNum")
    )$collect()
  
  return(result)
}
tic()
partitioned_parquet_polars_lazy()
toc()
```
:::

### Results for partitioned parquet files

```{r}
#| label: partitionned-parquet-results-benchmarking
#| message: false
#| warning: false

partitioned_parquet_bmk <- microbenchmark(
  "arrow (lazy) - from partitioned parquet file" = partitioned_parquet_arrow_lazy(),
  "dplyr (duckdb) - from partitioned parquet file" = partitioned_parquet_dplyr_duckdb(),
  "polars (lazy) - from partitioned parquet file" = partitioned_parquet_polars_lazy()$to_data_frame(),
  times = 5
 )
print(partitioned_parquet_bmk)
```
 
ðŸ‘‰ **Conclusion** of this little benchmark based **on partitioned parquet files**: as for unique parquet files, the big winner is `polars (lazy mode)` !  ðŸ†ðŸ†ðŸ† 