## From a csv file

For this comparison, we will use :

- For **polars (eager)**, the `pl$read_csv()` method
- For **polars (lazy)**, the `pl$scan_csv()` method
- For **R base**, the `read.csv()` method
- For **dplyr**, the `readr::read_csv()` method
- For **data.table**, the `data.table::fread()` method


::: {.panel-tabset}

## polars (eager)

```{r}
#| label: csv-eager-polars-benchmarking

csv_eager_polars <- function() {
# Reading the csv file (eager mode)
result_agg <- pl$read_csv(path = "Datasets/DataMultiTypes.csv")$
  # Conversion of 2 columns to Date format
  with_columns(
    pl$col("colDate1")$str$strptime(pl$Date, "%F %T", strict = FALSE),
    pl$col("colDate2")$str$strptime(pl$Date, "%F %T", strict = FALSE)
  )$
  # Creation of a diff column between 2 dates (in days)
  with_columns(
    (pl$col("colDate2") - pl$col("colDate1"))$dt$days()$alias("diff")
  )$
  # Filter rows
  filter(
    pl$col("colInt")>2000 & pl$col("colInt")<8000
  )$
  # Grouping and aggregation
  groupby(
    "colString")$
  agg(
    pl$col("colInt")$min()$alias("min_colInt"),
    pl$col("colInt")$mean()$alias("mean_colInt"),
    pl$col("colInt")$max()$alias("max_colInt"),
    pl$col("colNum")$min()$alias("min_colNum"),
    pl$col("colNum")$mean()$alias("mean_colNum"),
    pl$col("colNum")$max()$alias("max_colNum")
  )
  
  return(result_agg)
}
tic()
csv_eager_polars()
toc()
```

## polars (lazy)

```{r}
#| label: csv-lazy-polars-benchmarking
#| eval: false

csv_lazy_polars <- function() {
# Reading the csv file (eager mode)
result_agg <- pl$scan_csv(path = "Datasets/DataMultiTypes.csv")$
  # Conversion of 2 columns to Date format
  with_columns(
    pl$col("colDate1")$str$strptime(pl$Date, "%F %T", strict = FALSE),
    pl$col("colDate2")$str$strptime(pl$Date, "%F %T", strict = FALSE)
  )$
  # Creation of a diff column between 2 dates (in days)
  with_columns(
    (pl$col("colDate2") - pl$col("colDate1"))$dt$days()$alias("diff")
  )$
  # Filter rows
  filter(
    pl$col("colInt")>2000 & pl$col("colInt")<8000
  )$
  # Grouping and aggregation
  groupby(
    "colString")$
  agg(
    pl$col("colInt")$min()$alias("min_colInt"),
    pl$col("colInt")$mean()$alias("mean_colInt"),
    pl$col("colInt")$max()$alias("max_colInt"),
    pl$col("colNum")$min()$alias("min_colNum"),
    pl$col("colNum")$mean()$alias("mean_colNum"),
    pl$col("colNum")$max()$alias("max_colNum")
  )
  
  return(result_agg)
}
tic()
csv_lazy_polars()$collect()
toc()
```

## R base

```{r}
#| label: csv-rbase-benchmarking

csv_rbase <- function() {
  
  # Reading the csv file
  result <- read.csv("Datasets/DataMultiTypes.csv")
  
  # Conversion of 2 columns to Date format
  result$colDate1 <- as.Date(result$colDate1)
  result$colDate2 <- as.Date(result$colDate2)
  
  # Creation of a diff column between 2 dates (in days)
  result$diff <- round(
    as.integer(
      difftime(
        result$colDate2,
        result$colDate1,
        units = "days")
      ),
    0)
  
  # Filter rows
  result <- result[result$colInt>2000 & result$colInt<8000,]
  
  # Grouping and aggregation
  result_agg <- aggregate(cbind(colInt, colNum) ~ colString, 
                          data = result, 
                          FUN = function(x) c(mean = mean(x), 
                                              min = min(x), 
                                              max = max(x)))
  
  return(result_agg)
}

tic()
res_rbase <- csv_rbase()
toc()
print(res_rbase)
```

## dplyr

```{r}
#| label: csv-dplyr-benchmarking

csv_dplyr <- function() {
  
  # Reading the csv file
  result <- readr::read_csv("Datasets/DataMultiTypes.csv", show_col_types = FALSE)
  
  # Conversion of 2 columns to Date format
  result <- result |>
    mutate(
      colDate1 = as.Date(colDate1),
      colDate2 = as.Date(colDate2)
    )
  
  # Creation of a diff column between 2 dates (in days)
  result <- result |> 
    mutate(diff = round(as.integer(difftime(colDate2, colDate1, units = "days")),0))
  
  # Filter rows
  result <- result |>
    filter(
      colInt>2000 & colInt<8000
      )
  
  # Grouping and aggregation
  result_agg <- result |>
    group_by(colString) |> 
    summarise(
      min_colInt = min(colInt),
      mean_colInt = mean(colInt),
      mas_colInt = max(colInt),
      min_colNum = min(colNum),
      mean_colNum = mean(colNum),
      max_colNum = max(colNum)
  )
  
  return(result_agg)
}

tic()
res_dplyr <- csv_dplyr()
toc()
print(res_dplyr)
```

## data.table

```{r}
#| label: csv-datatable-optimised-benchmarking
csv_dt <- function() {
  
  result_agg <- as.data.table(data.table::fread("Datasets/DataMultiTypes.csv"))[, `:=`(
    
  colDate1 = as.Date(colDate1),
  colDate2 = as.Date(colDate2),
  diff = as.integer(difftime(colDate2, colDate1, units = "days"))
  
)][colInt > 2000 & colInt < 8000, .(
  
  min_colInt = min(colInt),
  mean_colInt = mean(colInt),
  max_colInt = max(colInt),
  min_colNum = min(colNum),
  mean_colNum = mean(colNum),
  max_colNum = max(colNum)
  
), by = colString]
  
  return(result_agg)
}
tic()
csv_dt()
toc()
```
:::

::: {.callout-note}
**The data processing performed is not entirely equivalent, since it includes in addition:**  
- for `polars (lazy mode)`, conversion to data.frame R at the end of processing  
- for `data.table`, conversion to dt format at the start, then conversion to data.frame R at the end of processing  
:::

### Results eager vs lazy mode

```{r}
#| label: csv-results-eager-lazy-benchmarking
#| message: false
#| warning: false

csv_bmk <- microbenchmark(
  "polars (eager) from csv file" = csv_eager_polars(),
  #"polars (lazy) from csv file" = csv_lazy_polars()$collect(),
  "R base - from csv file" = csv_rbase(),
  "dplyr - from csv file" = csv_dplyr(),
  "data.table - from csv file" = as.data.frame(csv_dt()),
  times = 5
 )
csv_bmk
```

ðŸ‘‰ The data processing is at least twice as fast with **lazy mode** `polars`! ðŸ†
