## From a csv file

::: {.panel-tabset}

## polars (eager)

```{r}
#| label: csv-polars-benchmarking

csv_polars <- function(variables) {
  
  # Reading the csv file (eager mode)
  result_pl <- pl$read_csv(path = "Datasets/DataMultiTypes.csv")
  
  # Conversion of 2 columns to Date format
  result_pl <- result_pl$with_columns(
    pl$col("colDate1")$str$strptime(pl$Date, "%F %T", strict = FALSE),
    pl$col("colDate2")$str$strptime(pl$Date, "%F %T", strict = FALSE)
  )
  
  # Creation of a diff column between 2 dates (in days)
  result_pl <- result_pl$with_columns(
    (pl$col("colDate2") - pl$col("colDate1"))$dt$days()$alias("diff")
  )
  
  return(result_pl)
}

tic()
res_polars <- csv_polars()
toc()
```

## R base

```{r}
#| label: csv-rbase-benchmarking

csv_rbase <- function() {
  
  # Reading the csv file
  result <- read.csv("Datasets/DataMultiTypes.csv")
  
  # Conversion of 2 columns to Date format
  result$colDate1 <- as.Date(result$colDate1)
  result$colDate2 <- as.Date(result$colDate2)
  
  # Creation of a diff column between 2 dates (in days)
  result$diff <- round(
    as.integer(
      difftime(
        result$colDate2,
        result$colDate1,
        units = "days")
      ),
    0)
  
  return(result)
}

tic()
res_rbase <- csv_rbase()
toc()
```

## dplyr

```{r}
#| label: csv-dplyr-benchmarking

csv_dplyr <- function() {
  
  # Reading the csv file
  result <- read.csv("Datasets/DataMultiTypes.csv")
  
  # Conversion of 2 columns to Date format
  result <- result |>
    mutate(
      colDate1 = as.Date(colDate1),
      colDate2 = as.Date(colDate2)
    )
  
  # Creation of a diff column between 2 dates (in days)
  result <- result |> 
    mutate(diff = round(as.integer(difftime(colDate2, colDate1, units = "days")),0))
  
  return(result)
}

tic()
res_dplyr <- csv_dplyr()
toc()
```

## data.table

```{r}
#| label: csv-datatable-benchmarking

csv_dt <- function(variables) {
  
  # Reading the csv file
  result <- read.csv("Datasets/DataMultiTypes.csv")
  
  # Conversion data.frame to data.table
  result_dt <- as.data.table(result)
  
  # Conversion of 2 columns to Date format
  result_dt <- result_dt[, `:=`(colDate1 = as.Date(colDate1),colDate2 = as.Date(colDate2))]
  
  # Creation of a diff column between 2 dates (in days)
  result_dt <- result_dt[, diff := as.integer(difftime(colDate2, colDate1, units = "days"))]
  
  return(result_dt)
}

tic()
res_dt <- csv_dt()
toc()
```


### Results

```{r}
#| label: csv-results-benchmarking
#| message: false
#| warning: false

microbenchmark(
  "R base" = csv_rbase(),
  "dplyr" = csv_dplyr(),
  "data.table" = as.data.frame(csv_dt()),
  "polars" = csv_polars()$to_data_frame(),
  times = 5,
  units = "s"
 )
```
 
 ::: {.callout-note title="2 comments on these results"}
The data processing performed is not entirely equivalent, since it includes :
- for `polars`, conversion to data.frame R at the end of processing
- for `data.table`, conversion to dt format at the start, then conversion to data.frame R at the end of processing
:::

ðŸ‘‰ The data processing is 56X faster with `polars`! ðŸ†