::: {.callout-note}
This chapter only deals with the **lazy execution** of Polars. It does not include a comparison with R base, dplyr and data.table
:::

## Introduction to lazy mode

Polars supports two modes of operation: **lazy** and **eager**.

Let's start this chapter by citing the official documentation:

> In the eager API the query is executed immediately while in the lazy API the query is only evaluated once it is 'needed'. Deferring the execution to the last minute can have significant performance advantages that is why the Lazy API is preferred in most cases. Delaying execution until the last possible moment allows Polars to apply automatic optimization to every query.

As you can see, with **lazy mode**, you give the engine the chance to analyse what you want to do in order to propose optimal execution (for both reading and transforming datasets). **Lazy evaluation** is a fairly common method of improving processing speed and is used by [Spark](https://spark.apache.org/docs/latest/index.html), among others.

So far in this book, we have only used the **eager mode** but fortunately all the syntax we've seen applies to **lazy mode** too. Whatever mode is used, queries will be executed transparently for users.

### Creation of a `LazyFrame` with lazy()

To convert a `DataFrame` to a `LazyFrame` we can use the `lazy()` contructor.

```{r}
#| label: creation-lazyFrame
as_polars_df(iris)$lazy()
```
We are no longer working on a `DataFrame` but on a `LazyFrame`.  

### First query passed on `LazyFrame`

Let's look at what happens when we request this `LazyFrame`:

```{r}
#| label: first-query-lazy
myquery <- as_polars_df(iris)$lazy()$filter(
  pl$col("Species") == "setosa"
)$select(
  pl$col("Species", "Petal.Length")
)
myquery
```

This way, we can display the **naive plan** (which means it is an non-optimized plan). Let's see what it contains for our example:  

- `FILTER [(col("Species")) == (Utf8(setosa))] FROM DF ["Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width"]` means that **once the entire datasets has been read into memory**, this `DataFrame` will be filtered for rows with Species equals to "setosa";   
- `PROJECT */5 COLUMNS` selects all 5 of the columns (* is the wildcard meaning all);  
- `SELECTION: "None"` means no rows will be filtered out.  

As indicated in the console, we can use the `explain()` method to see the **optimized plan**.

```{r}
#| label: optimized-query-lazy
myquery$explain()
```

This example shows a simple but surprisingly effective element of query optimisation: **projection**.   

Let's see what changed in this **optimized plan**:  

- `PROJECT 2/5 COLUMNS` selects only 2 columns;  
- `SELECTION: "[(col(\"Species\")) == (Utf8(setosa))]` means that Polars will apply the filter conditions on the `Species` column **as the datasets is being read.**  


We can see that Polars has identified that only 2 columns are needed to execute our query which is **memory efficient**! And Polars did this without me having to select these variables myself (for example with a `select` method).  
**The added value of Polars is that it applies some optimizations that I/you might not even have known about. 💪**

### Execute the plan

To actually execute the plan, we just need to invoke the **collect()** method. 

```{r}
#| label: collect-query-lazy
myquery$collect()
```