## `Lazy` vs `eager` mode comparison

In this first example we use the **eager API**:

```{r}
#| label: eager-query-steps
df <- pl$read_csv("examples/iris.csv")
df_small = df$filter(pl$col("Petal.Length") > 5)
df_agg = df_small$groupby("Species")$agg(pl$col("Petal.Width")$median())
df_agg
```

This example:

- Read the iris dataset.
- Filter the dataset based on Petal.Length
- Calculate the median of the Petal.Width per Species

**Every step is executed immediately returning the intermediate results**. This can be very **wastefull** as we might do work or load extra data that is not being used.  
If we instead used the **lazy API** and waited on execution untill all the steps are defined then the query planner could perform various optimizations. In this case:

- `Predicate pushdown`: Apply filters as early as possible while reading the dataset, thus only reading rows with sepal length greater than 5.
- `Projection pushdown`: Select only the columns that are needed while reading the dataset, thus removing the need to load additional columns

::: callout-tip
To consult the list of **optimisations** made by `Polars` on queries in **lazy mode**, see [this page](https://pola-rs.github.io/polars-book/user-guide/lazy/optimizations/)..
:::

Here is the equivalent code using the lazy API. At the end of the query, don't forget to use the `collect()` method to inform Polars that you want to execute it.

```{r}
#| label: lazy-query-steps
pl$read_csv("examples/iris.csv", lazy = TRUE)$
  filter(
    pl$col("Petal.Length") > 5)$
  groupby("Species")$
  agg(pl$col("Petal.Width")$median())$
  collect() # <- don't forget collect() here!
```

::: {.callout-important}
Use **lazy execution** will signficantly lower the load on memory & CPU thus allowing you to fit bigger datasets in memory and process faster.
:::

The next part will demonstrate this time saving. 👇
