## `Lazy` vs `eager` mode : fight! ⚔️

For this fight, we're going to use a fake dataset with 1 000 000 rows and 25 columns created with the {fakir} package. The code for creating this dataset is available at the beginning of this document.

This fight will take place over 3 rounds :

1. With an eager query versus a lazy query **from a DataFrame**
2. With an eager query versus a lazy query **from a csv file**
3. With an eager query versus a lazy query **from a parquet file**

### From a DataFrame

For this first round and as seen above, let's start with a simple query from a `DataFrame`:

```{r}
tic()
#| label: fight-eager_dataframe
pl$DataFrame(fake_data)$select(
    pl$col(c("region","departement","priority")) 
  )$
  filter(
    pl$col("region") == "Aquitaine")
toc()
```

As seen above, we're going to use the `lazy()` method to convert a `DataFrame` to a `LazyFrame`:

```{r}
tic()
#| label: fight-lazy_lazyframe
pl$DataFrame(fake_data)$lazy()$
  select(
    pl$col(c("region","departement","priority")) 
  )$
  filter(
    pl$col("region") == "Aquitaine")$
  collect() # don't forget collect() here!
toc()
```

### From a csv file

Now, the `eager` mode is represented here by the `read_csv()` method...

```{r}
tic()
#| label: fight-eager_read_csv
pl$read_csv("Datasets/fakir_file.csv", infer_schema_length=0)$ 
  select(
    pl$col(c("region","departement","priority","age")))$
  with_columns(
    pl$col("age")$cast(pl$Int32,strict = FALSE))$
  filter(
    pl$col("region") == "Bretagne")$
  groupby("departement","priority")$
  agg(pl$col("age")$mean())
toc()
```

... while the `lazy` method is represented by the `pl$scan_csv()`:

```{r}
#| label: fight-lazy_csv_reader
#| eval: false

tic()
pl$scan_csv("Datasets/fakir_file.csv", infer_schema_length=0)$
  select(
    pl$col(c("region","departement","priority","age")))$
  with_columns(
    pl$col("age")$cast(pl$Int32,strict = FALSE))$
  filter(
    pl$col("region") == "Bretagne")$
  groupby("departement","priority")$
  agg(pl$col("age")$mean())$
  collect()
toc()
```

We can clearly see that we save a lot of time when executing the `lazy` version of the code!


### From a parquet file

The `read_parquet()` method has not been implemented in the R Polars package, but for this fight we will use `arrow::read_parquet()` and {dplyr} syntax, which will compete with `pl$scan_parquet()`.


```{r}
#| label: fight-eager-read-parquet
#| message: false
tic()
arrow::read_parquet("Datasets/fakir_file.parquet", as_data_frame = FALSE) |>
  filter(region == "Bretagne") |> 
  group_by(departement,priority) |> 
  summarise(mymean=mean(age, na.rm = TRUE)) |> 
  arrange(departement) |>
  collect()
toc()
```

```{r}
#| label: fight-lazy-scan-parquet
tic()
pl$scan_parquet("Datasets/fakir_file.parquet")$
  filter( 
    pl$col("region") == "Bretagne")$
  groupby(c("departement","priority"))$
  agg(
    pl$col(c("age"))$mean()
)$sort("departement")$
  collect()
toc()
```
And it's another victory for the `lazy` execution!

::: {.callout-important}

Note that the {arrow} package also have ability to scan parquet files in a lazy way with the `arrow::open_dataset` function.

```{r}
#| label: fight-lazy-open-dataset
#| message: false
tic()
arrow::open_dataset("Datasets/fakir_file.parquet") |> 
  filter(region == "Bretagne") |>
  group_by(departement,priority) |> 
  summarise(mymean=mean(age, na.rm = TRUE)) |>
  arrange(departement) |>
  collect()
toc()
```

:::
