## Aggregation by group

Another frequently used data manipulation is the aggregation of data by group. To do this, we indicate in the `group_by()` method which column will be used to group the data.frame. And the `agg()` method which specifies the expression to aggregate.

The methods available for the `agg()` method are (in each group):

 - `first()` get the first element
 - `last()` get the last element
 - `n_unique()` get the number of unique elements
 - `count()` get the number of elements
 - `sum()` sum the elements
 - `min()` get the smallest element
 - `max()` get the largest element
 - `mean()` get the average of elements
 - `median()` get the median
 - `quantile()` calculate quantiles

 Here's a minimal example with `sum` and `mean` applied to 2 different columns:

::: {.panel-tabset}
## polars

```{r}
#| label: agg-polars
as_polars_df(iris)$
  group_by("Species")$
  agg(pl$sum("Petal.Length"),pl$col("Petal.Width")$mean())
```

## R base

```{r}
#| label: aggregate-cbind-rbase
aggregate(Petal.Length ~ Species, data = iris, FUN = sum)
aggregate(Petal.Width ~ Species, data = iris, FUN = mean)
```

## dplyr

```{r}
#| label: summarise-dplyr
data(iris)
iris %>%
  group_by(Species) %>%
  summarise(
    Petal.Length_sum = sum(Petal.Length),
    Petal.Width_mean = mean(Petal.Width)
  )
```

## data.table

```{r}
#| label: summarise-datatable
iris_dt <- as.data.table(iris)
iris_dt[, .(Petal.Length = sum(Petal.Length), Petal.Width = mean(Petal.Width)), by = Species]
```
:::

::: {.callout-tip}
Be careful! **Calling multiple aggregations on the same column produces columns of the same name** which generates an error with R.

You can use the `alias()` or `suffix()` method to ensure column names are unique.
For example:

```{r}
#| label: example-agg-with-alias

as_polars_df(iris)$
  group_by("Species")$
  agg(
    # With alias()
    pl$col(c("Petal.Length"))$sum()$alias("Petal.Length_Sum"),
    pl$col(c("Petal.Length"))$mean()$alias("Petal.Length_Mean"),
    # With suffix()
    pl$col(c("Petal.Width"))$sum()$name$suffix("_Sum"),
    pl$col(c("Petal.Width"))$mean()$name$suffix("_Mean")
  )
```
:::

You can perform more advanced aggregations, for example **calculate the sum of all the floating point columns** in iris:

::: {.panel-tabset}
## polars

```{r}
#| label: agg-advanced-polars
as_polars_df(iris)$
  group_by("Species")$
  agg(pl$col(pl$Float64)$sum())
```

## R base

```{r}
#| label: aggregate-advanced-rbase
iris_sum_by_species_base <- aggregate(iris[, sapply(iris, is.numeric)], by = list(Species = iris$Species), FUN = sum)
iris_sum_by_species_base
```

## dplyr

```{r}
#| label: summarise-advanced-dplyr
iris |>
  group_by(Species) %>%
  summarise(across(where(is.numeric), sum))
```

## data.table

```{r}
#| label: summarise-advanced-datatable
iris_dt <- as.data.table(iris)
iris_sum_by_species_dt <- iris_dt[, lapply(.SD, sum), by = Species, .SDcols = sapply(iris_dt, is.numeric)]
iris_sum_by_species_dt
```
:::