## Concatenate DataFrames

To perform concatenations, the `concat()` method must be used.  
In this section, we will see **vertical**, **horizontal** and **diagonal** concatenation.

Before we need to create some datasets to run our examples. Click to expand it! 👇

```{r}
#| label: creation-examples-datasets-concat
#| code-fold: true
#| results: 'hide'

dfup <- pl$DataFrame(
  col1 = c("a"),
  col2 = c("b"),
  col3 = c("c")
)
dfdown <- pl$DataFrame(
  col1 = c("x"),
  col2 = c("y"),
  col3 = c("z")
)
dfleft <- pl$DataFrame(col1 = c("a","b","c"))
dfright <- pl$DataFrame(col2 = c("x","y","z"))

dfup_df <- as.data.frame(dfup)
dfdown_df <- as.data.frame(dfdown)
dfleft_df <- as.data.frame(dfleft)
dfright_df <- as.data.frame(dfright)

dfup_dt <- as.data.table(dfup)
dfdown_dt <- as.data.table(dfdown)
dfleft_dt <- as.data.table(dfleft)
dfright_dt <- as.data.table(dfright)
```

### Vertical concatenation

To concatenate multiple `DataFrames` vertically (=by row), you can use the `concat()` method and the argument `how = "vertical"`.

::: {.panel-tabset}
## polars

```{r}
#| label: vertical-concatenation-polars
pl$concat(dfup,dfdown,how = "vertical")
```

## tidypolars

```{r}
#| label: vertical-concatenation-tidypolars
bind_rows_polars(dfup,dfdown)
```

## R base

```{r}
#| label: vertical-concatenation-rbase
rbind(dfup_df,dfdown_df)
```

## dplyr

```{r}
#| label: vertical-concatenation-dplyr
bind_rows(list(dfup_df,dfdown_df))
```

## data.table

```{r}
#| label: vertical-concatenation-datatable
rbindlist(list(dfup_dt, dfdown_dt))
```
:::

### Horizontal concatenation

To concatenate multiple `DataFrames` horizontally (=by col), you can use the `concat()` method and the argument `how = "horizontal"`.

::: {.panel-tabset}
## polars

```{r}
#| label: horizontal-concatenation-polars
pl$concat(dfleft,dfright,how = "horizontal")
```

## tidypolars

```{r}
#| label: horizontal-concatenation-tidypolars
bind_cols_polars(dfleft,dfright)
```

## R base

```{r}
#| label: horizontal-concatenation-rbase
cbind(dfleft_df,dfright_df)
```

## dplyr

```{r}
#| label: horizontal-concatenation-dplyr
bind_cols(list(dfleft_df,dfright_df))
```

## data.table

```{r}
#| label: horizontal-concatenation-datatable
data.table(dfleft_dt, dfright_dt)
```
:::

### Diagonal concatenation

To concatenate multiple `DataFrames` diagonally, you can use the `concat()` method and the argument `how = "diagonal"`.  
Diagonal concatenation is useful when the column names are not identical in initial `DataFrames`. 

::: {.panel-tabset}
## polars

```{r}
#| label: diagonal-concatenation-polars

dfup <- dfup$rename(col3="col4")
pl$concat(dfup,dfdown,how = "diagonal")
```

## tidypolars

```{r}
#| label: diagonal-concatenation-tidypolars
bind_rows_polars(dfup,dfdown)
```

## R base

```{r}
#| label: diagonal-concatenation-rbase
dfup_df <- as.data.frame(dfup)
dfup_df[setdiff(names(dfdown_df), names(dfup_df))] <- NA
dfdown_df[setdiff(names(dfup_df), names(dfdown_df))] <- NA
rbind(dfup_df,dfdown_df)
```

## dplyr

```{r}
#| label: diagonal-concatenation-dplyr
dfup_df <- as.data.frame(dfup)
bind_rows(list(dfup_df,dfdown_df))
```

## data.table

```{r}
#| label: diagonal-concatenation-datatable
dfup_dt <- as.data.table(dfup)
rbindlist(list(dfup_dt, dfdown_dt),fill = TRUE)
```
:::
