## Create your methods

With `R` you can create your own method/function with `function()`.  
Let's try to create a R function to captue some `DataFrame` transformations.  

Our simple function:  
- Takes a `DataFrame` as an input (argument `data`)  
- Convert Categorical columns into Strings  
- Make all Strings columns uppercase  
- And filter only the third first rows


```{r}
#| label: function-creation-polars
fn_transformation <- function(data) {
  
  data$
    # Convert Integer columns into Float 
    with_columns(
      pl$col(pl$Int32)$cast(pl$Float64))$
    # Make all Strings columns uppercase
    with_columns(
      pl$col(pl$String)$str$to_uppercase())$
    # Filter only the third first rows
    head(3)
  
}
```

Let's apply our method to `as_polars_df(iris)`:  

```{r}
#| label: function-use1-polars
fn_transformation(as_polars_df(mydf))
```

Our method `fn_transformation()` can now be re-used multiple times for example on another datasets:  

```{r}
#| label: function-use2-polars

Newdf <- pl$DataFrame(
  col_categ = factor(c("a","b","c")),
  col_str = c("x","y","z"),
  col_num = 1:3
)
  
fn_transformation(Newdf)
```

Of course, in real life, we will create functions that are more complicated than our example.

