### On DataFrames

#### Get Series from DataFrame

The `to_series()` method can be used to get one column from `DataFrame` as `Series`.

::: {.panel-tabset}
## polars

```{r}
#| label: to-series-Series
pl$DataFrame(iris)$select(pl$col("Petal.Length"))$to_series()
```

## R base

```{r}
#| label: tovector-rbase
iris$Petal.Length
```

## dplyr

```{r}
#| label: pull-dplyr
iris |>
  pull(Petal.Length)
```

## data.table

```{r}
#| label: tovector-datatable
iris_dt[, c(Petal.Length)]
```
:::

#### Get a slice of a DataFrame

The `to_series()` method can be used to get a slice of a `DataFrame`.

::: callout-important
With Polars, numeric default is 0! Thus the equivalent to `slice(1,3)` with Polars will be `2:4` in R Base and data.table and `slice(1,4)` with dplyr.
:::

::: {.panel-tabset}
## polars

```{r}
#| label: slice-Series
pl$DataFrame(iris)$slice(1,3)
```

## R base

```{r}
#| label: slice-rbase
iris[2:4,]
```

## dplyr

```{r}
#| label: slice-dplyr
iris |>
  slice(2:4)
```

## data.table

```{r}
#| label: slice-datatable
iris_dt[2:4,]
```
:::

#### Sort a DataFrame

The `sort()` method can be used to sort a `DataFrame`.

::: {.panel-tabset}
## polars

```{r}
#| label: sort-dataframe-polars
# Sort by one column
pl$DataFrame(iris
  )$sort("Species"
  )$head(3)
# Sort by two columns
pl$DataFrame(iris
  )$sort(c("Species","Petal.Length")
  )$head(3)
# Sort by two columns one in a decreasing manner and the other in an increasing manner
pl$DataFrame(iris
  )$sort(c("Species","Petal.Length"), descending = c(TRUE,FALSE)
  )$head(3)
```

## R base

```{r}
#| label: sort-dataframe-rbase
# Sort by one column
iris[order(iris$Species),][1:3,]
# Sort by two columns
iris[order(iris$Species,iris$Petal.Length),][1:3,]
# Sort by two columns one in a decreasing manner and the other in an increasing manner
iris[order(rev(iris$Species),iris$Petal.Length),][1:3,]
```

## dplyr

```{r}
#| label: sort-dataframe-dplyr
# Sort by one column
iris |>
  arrange(Species) |>
  head(3)
# Sort by two columns
iris |>
  arrange(Species, Petal.Length) |>
  head(3)
# Sort by two columns one in a decreasing manner and the other in an increasing manner
iris |>
  arrange(desc(Species), Petal.Length) |>
  head(3)
```

## data.table

```{r}
#| label: sort-dataframe-datatable
# Sort by one column
iris_dt[order(Species)][1:3,]
# Sort by two columns
iris_dt[order(Species,Petal.Length)][1:3,]
# Sort by two columns one in a decreasing manner and the other in an increasing manner
iris_dt[order(-Species,Petal.Length)][1:3,]
```
:::

#### Get a structure from a DataFrame

The `to_struct()` method can be used to get a structure from a `DataFrame`.

```{r}
#| label: to_struct-dataframe-polars
mystruc <- mydf$to_struct()
mystruc
```

`to_struct()` returns a `Series` which can be converted to a R list with `to_r` method.

```{r}
#| label: to_struct-to-r
mylist <- mystruc$to_r()
mylist
```

mylist is now a list where each element is a column of the initial DataFrame.

With `to_frame()` and `unnest()` methods, we can reconstruct the original DataFrame:

```{r}
back_df <- mystruc$to_frame()$unnest()
back_df
```

#### Drop all rows that contain null values

The `drop_nulls()` method can be used to drop all rows that contain null values in a `DataFrame`.

By default, `drop_nulls()` use all columns to drop rows:

::: {.panel-tabset}
## polars

```{r}
#| label: drop_nulls_all_columns-polars
data_pl <- pl$DataFrame(
  col1 = pl$Series(c(NA,"b","c")),
  col2 = pl$Series(c(1,2,NA))
)
data_pl$drop_nulls()
```

## R base
```{r}
#| label: drop_nulls_all_columns-rbase
data_df <- data.frame(
  col1 = c(NA, "b", "c"),
  col2 = c(1, 2, NA)
)
data_df[complete.cases(data_df), ]
```

## dplyr

```{r}
#| label: drop_nulls_all_columns-dplyr
data_df <- data.frame(
  col1 = c(NA, "b", "c"),
  col2 = c(1, 2, NA)
)
data_df %>%
  filter(complete.cases(.))
```

## data.table

```{r}
#| label: drop_nulls_all_columns-datatable
data_dt <- data.table(
  col1 = c(NA, "b", "c"),
  col2 = c(1, 2, NA)
)
na.omit(data_dt)
```
:::


If you want, you can specify a column (or multiple columns):

::: {.panel-tabset}
## polars

```{r}
#| label: drop_nulls_one_column-polars
data_pl$drop_nulls("col1")
```

## R base
```{r}
#| label: drop_nulls_one_column-rbase
data_df[!is.na(data_df$col1), ]
```

## dplyr
```{r}
#| label: drop_nulls_one_column-dplyr
data_df %>%
  filter(!is.na(col1))
```

## data.table
```{r}
#| label: drop_nulls_one_column-datatable
data_dt[complete.cases(data_dt[, .SD, .SDcols = "col1"]), ]
```
:::
