## Initial informations on a DataFrame

Here is a list of instructions that I frequently use to quickly get information about a `DataFrame`.

### Get the schema

A `DataFrame` has a `schema` attribute which is a named list of DataTypes.

```{r}
#| label: get-schema
mydf$schema
# This works also on LazyFrame
# mydf$lazy()$schema
```

### Get the dimensions

A `DataFrame` has a `shape` attribute which is a two length numeric vector of c(nrows,ncols).

```{r}
#| label: get-the-dimensions
mydf$shape
```

### Get columns types

A `DataFrame` has a `dtypes` attribute which is a list of dtypes (for data type) for every column of the `DataFrame`.  

Alternatively, the `dtype_strings()` method can be used to get columns types in a character/string vector.

```{r}
#| label: get-columns-types

# With dtypes attribute (with a "s" and without parentheses)
mydf$dtypes
# This works also on LazyFrame
# mydf$lazy()$dtypes
# With dtype_strings() method (wihout a "s" and with parentheses)
# mydf$dtype_strings()
```

### Get a glimpse

You can access a dense preview of a `DataFrame` by using the `glimpse()` method.  f
The formatting is done one line per column, so wide `DataFrame` show nicely. Each line will show the column name, the dtypes attributes and the first few values.  

```{r}
#| label: glimpse-method

as_polars_df(iris)$glimpse()
```

### Get summary statistics

You can access some summary statistics from a `DataFrame` by using the `describe()` method.

```{r}
#| label: describe-method

as_polars_df(iris)$describe()
```

### Get the rows number

A `DataFrame` has a `width` attribute which returns the width of the `DataFrame`.  

```{r}
#| label: get-rows-number

# With width attribute
mydf$width
```

### Get the columns number

A `DataFrame` has a `height` attribute which returns the width of the `DataFrame`.  

```{r}
#| label: get-columns-number

# With height attribute
mydf$height
```

### Get the columns names

A `DataFrame` has a `columns` attribute which returns the columns names of the `DataFrame` in a character vector.

```{r}
#| label: get-columns-names
mydf$columns
# This works also on LazyFrame
# mydf$lazy()$columns
```

### Rename columns

A character vector can be passed to the `columns` attribute to rename the columns names of the `DataFrame`.

```{r}
#| label: renames-columns
mydf$columns <- c("colA","colB")
mydf$columns
```

### Get the size

The `estimated_size()` method can be used to get an estimation of the total allocated size (in Bytes) of a `DataFrame`.

```{r}
#| label: get-size
mydf$estimated_size()
```

### Get the first n rows

The `head()` method can be used to get the first n rows of a `DataFrame`.

```{r}
#| label: get-first-n-rows

# To get the 2 first rows
mydf$head(2)
```


### Count values in a DataFrame

The `value_counts()` method can be used to count values in a Series of a DataFrame. 
`value_counts()` works with a `Series`. It must therefore be supplied either with square brackets or with the `select()` method. See [here](#select-columns) to learn about it.

::: {.panel-tabset}
## polars

```{r}
#| label: value-count-polars

# 1st option with square brackets
as_polars_df(iris)[,c("Species")]$value_counts()
# 2nd option with select() method
as_polars_df(iris)$select(pl$col("Species"))$to_series()$value_counts()
```

## R base

```{r}
#| label: value-count-rbase
table(iris$Species)
```

## dplyr

```{r}
#| label: value-count-dplyr
iris |>
  count(Species)
```

## data.table

```{r}
#| label: value-count-datatable
iris_dt[, .N, by = Species]
```
:::


### Count NA over columns in a DataFrame

The `null_count()` method can be used to count NA values of a DataFrame.

::: {.panel-tabset}
## polars

```{r}
#| label: null-count-polars
mydfNA <- pl$DataFrame(
  colA = as_polars_series(c("a",NA,"c")),
  colB = as_polars_series(c("d",NA,NA)))
mydfNA$null_count()
```

## R base

```{r}
#| label: sapply-NA-rbase
mydfNA <- data.frame(
  colA = c("a",NA,"c"),
  colB = c("d",NA,NA))
sapply(mydfNA, function(x) sum(is.na(x)))
```

## dplyr

```{r}
#| label: count-NA-dplyr
mydfNA |>
  summarise(across(everything(), ~sum(is.na(.))))
```

## data.table

```{r}
#| label: count-NA-datatable
mydfNA_dt <- as.data.table(mydfNA)
mydfNA_dt[, lapply(.SD, function(x) sum(is.na(x))), .SDcols = names(mydfNA_dt)]
```
:::
