## Dealing with missing values

We have already introduced missing values in [here](#missing-values). In this section, we will go further and understand how to deal with missing values with polars and R.

### Check if Series has missing values

The `is_null()` and `is_not_null()` methods can be used to check if `Series` has missing values.\
Theses methods are the equivalent of `is.na()` and `!is.na()` of R base.

Let's see two examples combining thes methods with `select()` and `filter()` methods:

::: {.panel-tabset}
## polars

```{r}
#| label: has-missing-values-polars
mydfNA <- pl$DataFrame(
  colA = c("a",NA,"c"),
  colB = c("d",NA,NA)
)

# Find null values
mydfNA$
  select(
    pl$col("colA"),
    pl$col("colA")$is_null()$alias("is_null"),
    pl$col("colA")$is_not_null()$alias("is_not_null")
  )

# Filter by null values
mydfNA$
  filter(
    pl$col("colA")$is_not_null() & pl$col("colB")$is_not_null()
  )
```

## tidypolars

```{r}
#| label: has-missing-values-tidypolars

# Find null values
mydfNA |> 
  select(
    colA) |> 
  mutate(
    is_null = is.na(colA),
    is_not_null = !is.na(colA)
  )

# Filter by null values
mydfNA |> 
  filter(!is.na(colA) & !is.na(colB))
```

## R base

```{r}
#| label: has-missing-values-rbase
mydfNA2 <- as.data.frame(mydfNA)

# Find null values
data.frame(
  colA = mydfNA2$colA,
  is_null = is.na(mydfNA2$colA),
  is_not_null = !is.na(mydfNA2$colA)
)

# Filter by null values
mydfNA2[!is.na(mydfNA2$colA) & !is.na(mydfNA2$colB), ]
```

## tidyr

```{r}
#| label: has-missing-values-tidypolars
mydfNA2 <- as.data.frame(mydfNA)

# Find null values
mydfNA2 |> 
  select(
    colA) |> 
  mutate(
    is_null = is.na(colA),
    is_not_null = !is.na(colA)
  )

# Filter by null values
mydfNA2 |> 
  filter(!is.na(colA) & !is.na(colB))
```
:::

### Replace missing values with a constant

The `fill_null()` method can be used to replace missing values in a `Series` with a constant.

Here's some examples where I replace missing values from all columns of a `DataFrame`:

::: {.panel-tabset}
## polars

```{r}
#| label: replace-missing-values-polars

mydfNA <- pl$DataFrame(
  colA = c("a",NA,"c"),
  colB = c("d",NA,NA)
)

# In the same columns
mydfNA$
  with_columns(
  pl$all()$fill_null("missing")
)

# In new columns suffixed by "_corrected"
mydfNA$
  with_columns(
  pl$all()$fill_null("missing")$name$suffix("_corrected")
)
```

## R base

```{r}
#| label: replace-missing-values-rbase

mydfNA2 <- as.data.frame(mydfNA)

# In the same columns
mydfNA2[is.na(mydfNA2)] <- "missing"
mydfNA2

# In new columns suffixed by "_corrected"
mydfNA2 <- as.data.frame(mydfNA)

transform(mydfNA2,
          colA_corrected = ifelse(is.na(colA), "missing", colA),
          colB_corrected = ifelse(is.na(colB), "missing", colB))

```

## dplyr

```{r}
#| label: replace-missing-values-dplyr

mydfNA2 <- as.data.frame(mydfNA)

# In the same columns
mydfNA2 %>%
  mutate(across(everything(), ~ifelse(is.na(.), "missing", .)))

# In new columns suffixed by "_corrected"
mydfNA2 %>%
  mutate(across(c(colA, colB), ~ifelse(is.na(.), "missing", .), .names = "{col}_corrected"))
```

## data.table

```{r}
#| label: replace-missing-values-datatable

mydfNA2 <- as.data.frame(mydfNA)
mydfNA2_dt <- as.data.table(mydfNA2)

# In the same columns
mydfNA2_dt[is.na(mydfNA2_dt)] <- "missing"
mydfNA2_dt

# In new columns suffixed by "_corrected"
mydfNA2_dt[,
           c("colA_corrected", "colB_corrected") := lapply(.SD, function(x) ifelse(is.na(x), "missing", x)),
           .SDcols = c("colA", "colB")]
mydfNA2_dt
```
:::

::: {.callout-important}
Be careful, **the `fill_null()` method can in some cases modify data types like `cast()`.**
This can happen, for example, when you're working on an integer column and you want to replace the missing values with a string => the column will then has a string dtype!
:::

### Replace missing values with a strategy

The `fill_null()` method of polars has a `strategy` argument for replacing missing values:

- `forward`: replace with the previous non-null value in the `Series`
- `backward`: replace with the next non-null value in the `Series`
- `min`: replace with the smallest value in the `Series`
- `max`: replace with the largest value in the `Series`
- `mean`: replace with the median value in the `Series`
- `zero`: replace with `0`
- `one`: replace with `1`

::: {.callout-note}
We can set a limit on how many rows to fill-forward or backward with `limit`
:::

Here's some examples :

```{r}
#| label: replace-missing-values-strategy-polars

# In the same columns
mydfNA <- pl$DataFrame(
  colA = c("a",NA,"c"),
  colB = c("d",NA,NA),
  colC = c(1,NA,3)
)
# With forward strategy
mydfNA$
  with_columns(
    pl$all()$fill_null(strategy = "forward")$name$suffix("_corrected")
  )
# With forward strategy and a limit
mydfNA$
  with_columns(
    pl$all()$fill_null(strategy = "forward", limit = 1)$name$suffix("_corrected")
  )
# With backward strategy
mydfNA$
  with_columns(
    pl$all()$fill_null(strategy = "backward")$name$suffix("_corrected")
  )
# With mean strategy only on "colC" column
mydfNA$
  with_columns(
    pl$col("colC")$fill_null(strategy = "mean")$name$suffix("_corrected")
  )
```

### Replace missing values with an expression

Of course, you are not limited to the built-in strategies of polars => with `fill_null()` you can also use expressions to replace missing values. It works with expression from the same column and from another column.

Here's some examples:

::: {.panel-tabset}
## polars

```{r}
#| label: replace-missing-values-expression-polars

# Replace missing values with the mean of the non-null values for that column
mydfNA$
  with_columns(
    pl$col("colC")$fill_null(strategy = "mean")$name$suffix("_corrected")
  )
# Replace missing values with the values from another column
mydfNA$
  with_columns(
    pl$col("colB")$fill_null(pl$col("colA"))$name$suffix("_corrected")
  )
```

## R base

```{r}
#| label: replace-missing-values-expression-rbase
mydfNA2 <- as.data.frame(mydfNA)

# Replace missing values with the mean of the non-null values for that column
mydfNA2$colC_corrected <- ifelse(is.na(mydfNA2$colC), mean(mydfNA2$colC, na.rm = TRUE), mydfNA2$colC)

mydfNA2 <- as.data.frame(mydfNA)
# Replace missing values with the values from another column
mydfNA2$colB_corrected <- ifelse(is.na(mydfNA2$colB), mydfNA2$colA, mydfNA2$colB)
mydfNA2
```

## dplyr

```{r}
#| label: replace-missing-values-expression-dplyr
mydfNA2 <- as.data.frame(mydfNA)

# Replace missing values with the mean of the non-null values for that column
mydfNA2 %>%
  mutate(colC_corrected = ifelse(is.na(colC), mean(mydfNA2$colC, na.rm = TRUE), colC))
# Replace missing values with the values from another column
mydfNA2 %>%
  mutate(colB_corrected = ifelse(is.na(colB), colA, colB))
```

## data.table

```{r}
#| label: replace-missing-values-expression-datatable

mydfNA2 <- as.data.frame(mydfNA)
mydfNA2_dt <- as.data.table(mydfNA2)

# Replace missing values with the mean of the non-null values for that column
mydfNA2_dt[, colC_corrected := ifelse(is.na(colC), mean(mydfNA2_dt$colC, na.rm = TRUE), colC)]
mydfNA2_dt

# Replace missing values with the values from another column
mydfNA2_dt[, colB_corrected := ifelse(is.na(colB), colA, colB)]
mydfNA2_dt
```
:::

### Replace missing values with a sequence of columns

The `coalesce()` method can be used to replace missing values based on a sequence of columns.

Here's an example creating a new column "col4" that has the first non-null value as we go through some columns (in order!):

::: {.panel-tabset}
## polars

```{r}
#| label: coalesce-polars

mynewdfNA <- pl$DataFrame(
  col1 = c(NA,"y",NA),
  col2 = c(NA,"v","w"),
  col3 = c("r","s",NA)
)

mynewdfNA$
  with_columns(
    pl$coalesce("col1","col2","col3")$alias("col4")
  )
```

## R base

```{r}
#| label: coalesce-rbase
mynewdfNA2 <-  as.data.frame(mynewdfNA)

mynewdfNA2$col4 <- coalesce(mynewdfNA2$col1, mynewdfNA2$col2, mynewdfNA2$col3)
mynewdfNA2
```

## dplyr

```{r}
#| label: coalesce-dplyr
mynewdfNA2 <-  as.data.frame(mynewdfNA)

mynewdfNA2 %>%
  mutate(col4 = coalesce(col1, col2, col3))
```

## data.table

```{r}
#| label: coalesce-datatable
mynewdfNA2 <-   as.data.frame(mynewdfNA)
mynewdfNA2_dt <- as.data.table(mynewdfNA2)

mynewdfNA2_dt[, col4 := fcoalesce(col1, col2, col3)]
mynewdfNA2_dt
```
:::
