## Pivot a DataFrame

### From long to wide

The `pivot()` method can be used to pivot a `DataFrame` from long to wide.  

Let's go for a first example :  

```{r}
#| label: creation-df-for-pivot-wider
df <- data.frame(
  country = c(rep("France",3),rep("Italy","3")),
  city = c("Paris", "Lille", "Nice", "Roma", "Milan", "Napoli"),
  location = c("North","North","South","South","North","South"),
  population = c(2.1, 0.2, 0.4, 2.8, 1.4, 3.0)
)
```


::: {.panel-tabset}
## polars

```{r}
#| label: first-wide-pivot-polars
pl$DataFrame(df)$pivot(
  index = "country", 
  columns = "city",
  values = "population", 
)
```

## R base
```{r}
#| label: first-wide-pivot-rbase
reshape(df[,-which(names(df) %in% c("location"))], 
        idvar = "country", 
        timevar = "city", 
        direction = "wide")
```

## tidyr

```{r}
#| label: first-wide-pivot-polars-tidyr
df |>
  pivot_wider(
    id_cols = country,
    names_from = city,
    values_from = population
  )
```

## data.table

```{r}
#| label: first-wide-pivot-datatable
df_dt <- as.data.table(df)
dcast(df_dt, country ~ city, value.var = "population")
```
:::

You can also aggregate the results using a function that you enter in the argument `aggregate_function` in `pivot()` method.  

In this case, the `aggregate_function` argument of `pivot()` is the equivalent of `values_fn` of `pivot_wider()` from `{tidyr}` and `fun.aggregate` of `dcast()` from `{data.table}`.

::: {.panel-tabset}
## polars

```{r}
#| label: mean-wide-pivot-polars
pl$DataFrame(df)$pivot(
  index = "country", 
  columns = "location",
  values = "population",
  aggregate_function = "mean"
)
```

## R base
```{r}
#| label: mean-wide-pivot-rbase
df_summary <- aggregate(population ~ country + location, data = df, FUN = mean)
df_final <- reshape(df_summary, idvar = "country", timevar = "location", direction = "wide")
colnames(df_final) <- c("country", "North", "South")
df_final
```

## tidyr

```{r}
#| label: mean-wide-pivot-polars-tidyr
df |>
  pivot_wider(id_cols = country,
              names_from = location, 
              values_from = population, 
              values_fn = mean)
```

## data.table

```{r}
#| label: mean-wide-pivot-datatable
df_dt <- as.data.table(df)
dcast(df_dt, country ~ location, value.var = "population", fun.aggregate = mean)
```
:::

However with `{polars}`, we can also run an expression as an aggregation function.  
With `{tidyr}` and `{data.table}`, you need to calculate this in advance.  
For example:

::: {.panel-tabset}
## polars

```{r}
#| label: expr-wide-pivot-polars
pl$DataFrame(df)$pivot(
  index = "country", 
  columns = "location",
  values = "population",
  aggregate_function = pl$element()$sum()$sqrt()
)
```

## R base
```{r}
#| label: expr-wide-pivot-rbase
df_summarized <- aggregate(population ~ country + location, df, FUN = function(x) sqrt(sum(x)))
df_final <- reshape(df_summarized, idvar = "country", timevar = "location", direction = "wide")
colnames(df_final) <- c("country", "North", "South")
df_final
```

## tidyr

```{r}
#| label: expr-wide-pivot-polars-tidyr
#| message: false
df |>
  group_by(country, location) |>
  summarise(population_sum = sqrt(sum(population))) |>
  pivot_wider(names_from = location, values_from = population_sum)
```

## data.table

```{r}
#| label: expr-wide-pivot-datatable
dt_final <- df_dt[, .(population_sum = sqrt(sum(population))), by = .(country, location)]
dcast(dt_final, country ~ location, value.var = "population_sum")
```
:::

### From wide to long

The `melt()` method can be used to pivot a `DataFrame` from wide to long.  

Let's see with an example :  

```{r}
#| label: creation-df-for-pivot-longer
df <- data.frame(
  country = c("France","Italy"),
  North = c(1.1,1.4),
  South = c(0.4,2.9)
)
```


::: {.panel-tabset}
## polars

```{r}
#| label: first-long-pivot-polars
pl$DataFrame(df)$melt(
  id_vars = "country",
  value_vars = c("North","South")
)
```

## R base
```{r}
#| label: first-long-pivot-rbase
melted_df  <- reshape(df, 
        varying = c("North", "South"), 
        v.names = "value", 
        idvar = "country", 
        times = c("North", "South"), 
        timevar = "variable", 
        direction = "long")
rownames(melted_df) <- NULL
melted_df
```

## tidyr

```{r}
#| label: first-long-pivot-polars-tidyr
df |>
  pivot_longer(
    !country,
    names_to = "variable",
    values_to = "value"
  )
```

## data.table

```{r}
#| label: first-long-pivot-datatable
df_dt <- as.data.table(df)
melt(df_dt, id.vars = "country", variable.name = "variable", value.name = "value")
```
:::