## Remove columns

### Removing by name

To remove columns from a DataFrame, the `drop` expression must be used.  
Here's an example with equivalent syntax:  

```{r}
#| label: reload-iris
#| echo: false
#| include: false

data(iris)
iris_dt <- as.data.table(iris)
```


::: {.panel-tabset}  
## polars

```{r}
#| label: remove-column-polars
pl$DataFrame(iris)$
  drop(c("Petal.Width","Sepal.Length","Sepal.Width"))$
  head(3)
```

## R base

```{r}
#| label: remove-column-rbase
iris[1:3,!(names(iris) %in% c("Petal.Width","Sepal.Length","Sepal.Width"))]
```

## dplyr

```{r}
#| label: remove-column-dplyr
iris |>
  select(-c(Petal.Width,Sepal.Length,Sepal.Width)) |>
  head(3)
```

## data.table

```{r}
#| label: remove-column-datatable
iris_dt[, c("Petal.Width","Sepal.Length","Sepal.Width") := NULL][1:3,]
```
:::

### Removing by name pattern

To remove columns by name pattern from a DataFrame, the `drop` expression must be used.  
Let's see an example where you want to drop columns whose names starts with "Petal" in `iris`.

::: {.panel-tabset}  
## polars

```{r}
#| label: remove-column-pattern-polars

num_to_drop <- !grepl("^Petal",pl$DataFrame(iris)$columns)
col_to_drop <- pl$DataFrame(iris)$columns[num_to_drop]

pl$DataFrame(iris)$
  drop(col_to_drop)$
  head(3)
```

## R base

```{r}
#| label: remove-column-pattern-rbase
iris[1:3,!grepl("^Petal",names(iris))]
```

## dplyr

```{r}
#| label: remove-column-pattern-dplyr
iris |>
  select(-starts_with("Petal")) |>
  head(3)
```

## data.table

```{r}
#| label: remove-column-pattern-datatable
#| warning: false
iris_dt[, which(grepl("^Petal$", names(iris_dt))):=NULL][1:3,]
```
:::