## Select columns

### Selecting by name

The first option for selecting columns of a DataFrame is to use square brackets `[]`.  
The second option is to use the `select()` method. In this case, it must be used with the `col()` method which allows to designate the columns to be selected with a character vector.

::: {.panel-tabset}
## polars

```{r}
#| label: select-polars

# 1st option : with square brackets syntax
as_polars_df(iris)[1:3,c("Petal.Length","Petal.Width")] 
# 2nd option : with select() method
as_polars_df(iris)$select(
  pl$col("Petal.Length","Petal.Width")
)$head(3) # display the first 3 lines
```

## tidypolars

```{r}
#| label: select-tidypolars

as_polars_df(iris) |>
  select(Petal.Length,Petal.Width) |>
  head(3)
```

## R base

```{r}
#| label: select-rbase
iris[1:3,c("Petal.Length","Petal.Width")] 
```

## dplyr

```{r}
#| label: select-dplyr
iris |>
  select(Petal.Length,Petal.Width) |>
  head(3)
```

## data.table

```{r}
#| label: select-datatable
iris_dt[1:3,.(Petal.Length,Petal.Width)]
```
:::

There are **two major reasons why you should use the `select()` method over the syntax with square brackets:**  
- When you select and transform multiple columns with `select()` method, Polars will run these selections **in parallel**;  
- Expressions used in `select()` method can be optimised in [lazy mode](#introduction-to-lazy-mode) by the query optimizer.

### Selecting by data type

To select columns by data type from a `DataFrame`, you can pass a data type to `pl$col` expression.
For example:

::: {.panel-tabset}
## polars

```{r}
#| label: select-datatype-polars

# Select only Float64 columns
as_polars_df(iris)$select(pl$col(pl$Float64))$head(3)
# Select only numeric columns
as_polars_df(iris)$select(cs$numeric())$head(3) 
```

## tidypolars

```{r}
#| label: select-datatype-tidypolars

# Select only numeric columns
as_polars_df(iris) %>% 
  select(where(is.numeric)) %>% 
  head(3)
```

## R base

```{r}
#| label: select-datatype-rbase

# Select only factor columns
as.data.frame(iris[1:3, sapply(iris, is.factor)])
# Select only numeric columns
iris[1:3, sapply(iris, is.numeric)]
```

## dplyr

```{r}
#| label: select-datatype-dplyr

# Select only factor columns
iris %>% 
  select_if(is.factor) %>%
  head(3)
# Select only numeric columns
iris %>% 
  select_if(is.numeric) %>%
  head(3)
```

## data.table

```{r}
#| label: select-datatype-datatable

# Select only factor columns
iris_dt[, .SD, .SDcols = is.factor][1:3]
# Select only numeric columns
iris_dt[, .SD, .SDcols = is.numeric][1:3]
```
:::

::: {.callout-note}
You can also select columns from a DataFrame using a name pattern. See [an example by removing a column](#removing-by-name-pattern)
:::

### Selecting with a regex

To select columns with a regex from a `DataFrame`, you can pass it in `pl$col` expression.
For example, to select all columns that starts with "Sepal" in iris dataset:  

::: {.panel-tabset}
## polars

```{r}
#| label: select-regex-polars

as_polars_df(iris)$select(
  pl$col("^Sepal.*$")
)$head(3) # display the first 3 lines

```

## tidypolars

```{r}
#| label: select-regex-tidypolars

as_polars_df(iris) %>%
  select(starts_with("Sepal")) %>%
  head(3)
```

## R base

```{r}
#| label: select-regex-rbase

iris[, grep("^Sepal.*$", colnames(iris))]
```

## dplyr

```{r}
#| label: select-regex-dplyr

iris %>%
  select(starts_with("Sepal")) %>%
  head(3)
```

## data.table

```{r}
#| label: select-regex-datatable

iris_dt[, .SD, .SDcols = grep("^Sepal", colnames(iris_dt))][1:3]
```
:::

### Selecting - other ways

The selectors API of polars enables to use other methods to select columns.  
Here's a few examples selecting the first and last column of a `DataFrame` respectively with `first()` and `last()` methods:  

```{r}
#| label: first-last-polars

# Select the first column
as_polars_df(iris)$select(
  pl$first()
)$head(3)

# Select the last column
as_polars_df(iris)$select(
  pl$last()
)$head(3)
```
