## Select columns

### Selecting by name

The first option for selecting columns of a DataFrame is to use square brackets `[]`.  
The second option is to use the `select()` method. In this case, it must be used with the `col()` method which allows to designate the columns to be selected with a character vector.

::: {.panel-tabset}
## polars

```{r}
#| label: select-polars

# 1st option : with square brackets syntax
pl$DataFrame(iris)[1:3,c("Petal.Length","Petal.Width")] 
# 2nd option : with select() method
pl$DataFrame(iris)$select(
  pl$col(c("Petal.Length","Petal.Width"))
)$head(3) # display the first 3 lines
```

## R base

```{r}
#| label: select-rbase
iris[1:3,c("Petal.Length","Petal.Width")] 
```

## dplyr

```{r}
#| label: select-dplyr
iris |>
  select(Petal.Length,Petal.Width) |>
  head(3)
```

## data.table

```{r}
#| label: select-datatable
iris_dt[1:3,.(Petal.Length,Petal.Width)]
```
:::

**Keep in mind that when:**  
- Square brackets are used:  
    - **Selecting only one column** from a `DataFrame`, the output is a `Series`;  
    - Whereas **if there is more than one column selected**, the output is a `DataFrame`.  
- `select()` method is used:  
    - **The output is always a `DataFrame`** rather than a `Series` even if one column is selected.  
    => If you need a `Series` you can use the `to_series()` method. See [here](#get-series-from-dataFrame).

```{r}
#| label: select-output-polars

inherits(pl$DataFrame(iris)[,"Petal.Length"],"Series")
inherits(pl$DataFrame(iris)[,"Petal.Length"],"DataFrame")
inherits(pl$DataFrame(iris)[,c("Petal.Length","Petal.Width")],"DataFrame")
inherits(pl$DataFrame(iris)[,c("Petal.Length","Petal.Width")],"Series")
```

With Polars if you want to obtain a result in an R data.frame, you can simply add the method `to_data_frame()` at the end of the method chaining. See [here](#from-seriesdataframes-to-vectordata.frames) for examples.

::: {.callout-tip}
Beyond the minor differences discussed above, there are **two major reasons why you should use the `select()` method over the syntax with square brackets:**  
- When you select and transform multiple columns with `select()` method, Polars will run these selections **in parallel**;  
- Expressions used in `select()` method can be optimised in [lazy mode](#introduction-to-lazy-mode) by the query optimizer.
:::

### Selecting by data type

To select columns by data type from a `DataFrame`, you can pass a data type to `pl$col` expression.
For example:

::: {.panel-tabset}
## polars

```{r}
#| label: select-datatype-polars

# Select only categorical columns
pl$DataFrame(iris)$select(pl$col(pl$Categorical))$head(3) 
# Select only Float64 columns
pl$DataFrame(iris)$select(pl$col(pl$Float64))$head(3)
```

## R base

```{r}
#| label: select-datatype-rbase

# Select only factor columns
as.data.frame(iris[1:3, sapply(iris, is.factor)])
# Select only numeric columns
iris[1:3, sapply(iris, is.numeric)]
```

## dplyr

```{r}
#| label: select-datatype-dplyr

# Select only factor columns
iris %>% 
  select_if(is.factor)
# Select only numeric columns
iris %>% 
  select_if(is.numeric)
```

## data.table

```{r}
#| label: select-datatype-datatable

# Select only factor columns
iris_dt[, .SD, .SDcols = is.factor]
# Select only numeric columns
iris_dt[, .SD, .SDcols = is.numeric]
```
:::

::: {.callout-note}
You can also select columns from a DataFrame using a name pattern. See [an example by removing a column](#removing-by-name-pattern)
:::

### Selecting with a regex

To select columns with a regex from a `DataFrame`, you can pass it in `pl$col` expression.
For example, to select all columns that starts with "Sepal" in iris dataset:  

::: {.panel-tabset}
## polars

```{r}
#| label: select-regex-polars

pl$DataFrame(iris)$select(
  pl$col("^Sepal.*$")
)$head(3) # display the first 3 lines

```

## R base

```{r}
#| label: select-regex-rbase

iris[, grep("^Sepal.*$", colnames(iris))]
```

## dplyr

```{r}
#| label: select-regex-dplyr

iris %>%
  select(starts_with("Sepal"))
```

## data.table

```{r}
#| label: select-regex-datatable

iris_dt[, .SD, .SDcols = grep("^Sepal", colnames(iris_dt))]
```
:::

### Selecting using lists

It is also possible to select columns from a `DataFrame` using `R lists` which can be very practical depending on the case. Here's an example:

```{r}
#| label: select-lists-polars
data(iris)

l_expr = list(
  pl$col("Petal.Length"),
  pl$col("Species")
)

# Select only categorical columns
pl$DataFrame(iris)$select(l_expr)$head(3) 
```
This way of selecting columns also works with simple R lists:

```{r}
#| label: select-simple-lists-polars
data(iris)

l_expr = list(
  "Petal.Length",
  "Species",
  "Petal.Width"
)

# Select only categorical columns
pl$DataFrame(iris)$select(l_expr)$head(3) 
```

### Removing columns

To remove one or multiple columns from a `DataFrame` you can use the `all()` and `exclude()` methods.   
Here's an example:

::: {.panel-tabset}
## polars

```{r}
#| label: remove-columns-polars

pl$DataFrame(iris)$select(
  pl$all()$exclude(c("Petal.Length","Species"))
)$head(3) # display the first 3 lines
```

## R base

```{r}
#| label: remove-columns-rbase
iris[1:3,-which(names(iris) %in% c("Petal.Length","Species"))]
```

## dplyr

```{r}
#| label: remove-columns-dplyr

# Select only factor columns
iris %>% 
  select(-Petal.Length,-Species)
```

## data.table

```{r}
#| label: remove-columns-datatable
res <- iris_dt[, c("Petal.Length","Species"):=NULL]
res
data(iris)
```
:::

::: {.callout-note}
All the ways we've seen in this section for selecting columns from a `DataFrame` (by name, data type, with a regex and with a list) also work with `pl$all()$exclude()`!
:::

