## Select columns

The first option for selecting columns of a DataFrame is to use square brackets `[]`.  
The second option is to use the `select()` method. In this case, it must be used with the `col()` method which allows to designate the columns to be selected.

::: {.panel-tabset}
## polars

```{r}
#| label: select-polars

# 1st option : with square brackets syntax
pl$DataFrame(iris)[1:3,c("Petal.Length","Petal.Width")] 
# 2nd option : with select() method
pl$DataFrame(iris)$select(
  pl$col(c("Petal.Length","Petal.Width"))
)$head(3) # display the first 3 lines
```

## R base

```{r}
#| label: select-rbase
iris[1:3,c("Petal.Length","Petal.Width")] 
```

## dplyr

```{r}
#| label: select-dplyr
iris |>
  select(Petal.Length,Petal.Width) |>
  head(3)
```

## data.table

```{r}
#| label: select-datatable
iris_dt[1:3,.(Petal.Length,Petal.Width)]
```
:::

::: {.callout-note}
With Polars if you want to obtain a result in an R data.frame, you can simply add the method `to_data_frame()` at the end of the method chaining.
:::
