## Others useful methods

### On Series

#### Change name of Series

The `alias()` method is very useful especially in method chaining operation.\
With R base, the syntax is longer.

::: {.panel-tabset}
## polars

```{r}
#| label: change-name-Series
pl$Series(1:3, name = "toto")$alias("titi")
```

## R base

```{r}
#| label: change-name-rbase
toto <- 1:3
titi <- toto
rm(toto)
```
:::

#### Reduce Boolean Series

The `all()` and `any()` methods can be used to check if all or any values in a vector evaluate to TRUE for some expression.

::: {.panel-tabset}
## polars

```{r}
#| label: all-Series
# all(pl$Series(c(TRUE,TRUE))) doesn't work
pl$Series(c(TRUE, TRUE, NA))$all()
pl$Series(c(TRUE, TRUE, FALSE))$all()
pl$Series(c(TRUE, TRUE, TRUE))$all()
```

## R base

```{r}
#| label: all-rbase
all(c(TRUE,TRUE,NA))
all(c(TRUE,TRUE,FALSE))
all(c(TRUE,TRUE,TRUE))
```
:::

#### Get data type of Series

The `dtype()` method can be used to get data type of `Series`.

::: {.panel-tabset}
## polars

```{r}
#| label: dtype-Series

pl$Series(letters)$dtype
pl$Series(c(1, 2))$dtype
pl$DataFrame(iris)$select("Species")$to_series()$dtype
```

## R base

```{r}
#| label: dtype-rbase
infer_type(letters)
infer_type(c(1, 2))
infer_type(iris$Species)
```
:::

::: {.callout-tip}
Polars is strongly typed. `print(ls(pl$dtypes))` returns the full list of valid Polars types. Caution, some type names differ from what they are called in R base. See below!
:::

::: {.panel-tabset}
## polars

```{r}
#| label: dtype-examples-Series
pl$Series(c("x","y","z"))$dtype
pl$Series(c(1, 2, 3))$dtype
pl$Series(c(1:3))$dtype
pl$Series(c(TRUE,FALSE))$dtype
pl$DataFrame(iris)$select("Species")$to_series()$dtype
pl$Series(Sys.Date())$dtype
pl$Series(c(0,1))$dtype
```

## R base

```{r}
#| label: dtype-examples-rbase
typeof(c("x","y","z"))
typeof(c(1, 2, 3))
typeof(c(1:3))
typeof(c(TRUE,FALSE))
typeof(iris$Species)
typeof(Sys.Date())
```
:::

To summarise the main types between Polars and R:

| Polars      | R Base    |
|-------------|-----------|
| Utf8        | character |
| Float64     | double    |
| Int32       | integer   |
| Boolean     | logical   |
| Categorical | Factor    |
| Date        | Date      |

To learn more about Data types in Polars, see [here](https://pola-rs.github.io/polars-book/user-guide/concepts/data-types/).

#### Cast

The `cast()` method can be used to convert the data types of a column to a new one.

::: {.panel-tabset}
## polars

```{r}
#| label: cast-polars
pl$DataFrame(iris)$with_columns(
  pl$col("Petal.Length")$cast(pl$Int8), # The "Petal.Length" column is converted into integers
  pl$col("Species")$cast(pl$Utf8) # The "Species" column is converted into strings
  )$schema
```

## R base

```{r}
#| label: cast-rbase
data(iris)
iris$Petal.Length <- as.integer(iris$Petal.Length)
iris$Species <- as.integer(iris$Species)
str(iris)
```

## dplyr

```{r}
#| label: cast-dplyr
data(iris)
iris |>
  mutate(
    Petal.Length = as.integer(Petal.Length),
    Species = as.character(Species)) |>
  str()
```

## data.table

```{r}
#| label: cast-datatable
iris_dt[, `:=`(Petal.Length = as.integer(Petal.Length),
               Species = as.character(Species))]
str(iris_dt)
```
:::

::: {.callout-tip}
When working with very large tables we can **Reduce the memory footprint** by modifying the number of bits allocated to an element. ⚠️
:::

For example, the example below illustrates how converting Float64 to Float8 reduces memory usage:

```{r}
pl$DataFrame(iris)$estimated_size()

#| label: reduce-footprint-polars
pl$DataFrame(iris)$with_columns(
  pl$col("Petal.Length")$cast(pl$Float32), 
  pl$col("Petal.Width")$cast(pl$Float32), 
  pl$col("Sepal.Length")$cast(pl$Float32), 
  pl$col("Sepal.Width")$cast(pl$Float32)
  )$estimated_size()
```

When performing downcasting, it is crucial to ensure that the chosen number of bits is sufficient to accommodate the largest and smallest numbers in the column.

A quick reminder:

| Type    | Range                      | Accuracy                |
|---------|----------------------------|-------------------------|
| Int8    | -128 to +127               |                         |
| Int16   | -32768 to +32767           |                         |
| Int32   | -2147483648 to +2147483647 |                         |
| Int64   | --2E63 to --2E63-1         |                         |
| Float32 | -3.4E+38 to +3.4E+38       | about 7 decimal digits  |
| Float64 | -1.7E+308 to +1.7E+308     | about 16 decimal digits |

#### Check if Series is numeric

The `is_numeric()` method can be used to check if `Series` is numeric.\
Note that unlike R base, there is no method to check if a Series is character (in this case, its type is anyway Utf8).

::: {.panel-tabset}
## polars

```{r}
#| label: isnumeric-Series
pl$Series(1:4)$is_numeric()
pl$Series(c("a", "b", "c"))$is_numeric()
```

## R base

```{r}
#| label: isnumeric-rbase
is.numeric(1:4)
is.numeric(c("a","b","c"))
```
:::

#### Check if Series is sorted

The `is_sorted()` method can be used to check if `Series` is sorted.\
Note that R base provides `is.unsorted()` which returns the opposite boolean to `is_sorted()` of Polars.

::: {.panel-tabset}
## polars

```{r}
#| label: issorted-Series
pl$Series(1:4)$is_sorted()
pl$Series(c(1,3,2))$is_sorted()
```

## R base

```{r}
#| label: isunsorted-rbase
is.unsorted(1:4)
is.unsorted(c(1,3,2))
```
:::

#### Get length of a Series

The `len()` method can be used to get the length of a `Series`.

::: {.panel-tabset}
## polars

```{r}
#| label: len-Series
pl$Series(1:4)$len()
```

## R base

```{r}
#| label: length-rbase
length(1:4)
```
:::

#### Check if Series are equal

The `series_equal()` method can be used to check if a `Series` is equal with another `Series`.

::: {.callout-tip}
Caution, if two series are identical but one is named and the other is not then `series_equal()` returns FALSE.
:::

::: {.panel-tabset}
## polars

```{r}
#| label: series-equal-Series
pl$Series(1:4)$series_equal(pl$Series(1:4))
pl$Series(1:4,name = "toto")$series_equal(pl$Series(1:4))
```

## R base

```{r}
#| label: identical-rbase
identical(1:4,1:4)
```
:::

#### Convert Series to Polars DataFrame

The `to_frame()` method can be used to convert a `Series` to a `DataFrame`.\
In this case, a `DataFrame` with only one column will be created. If the `Series` is initially named then the column of the `DataFrame` will be named as such.

```{r}
#| label: to-frame-Series
pl$Series(1:3, "toto")$to_frame()
```

#### Get value Counts of a Series

The `value_counts()` method can be used to get a value counts of a `Series`.

::: {.panel-tabset}
## polars

```{r}
#| label: value-counts-Series
pl$Series(iris$Species)$value_counts()
```

## R base

```{r}
#| label: table-rbase
table(iris$Species)
```

## dplyr

```{r}
#| label: count-dplyr
iris |>
  count(Species)
```

## data.table

```{r}
#| label: value-counts-datatable
iris_dt[, .N, by = Species]
```
:::

#### Sum across `Series`

The `sum()` method can be used to get a sum of a `Series`.

- From a single `Series`:

::: {.panel-tabset}
## polars

```{r}
#| label: first-sum-polars
pl$Series(1:3)$sum()
```

## R base

```{r}
#| label: first-sum-rbase
sum(c(1:3))
```
:::


- From a `DataFrame` and a column as **a string**:

::: {.panel-tabset}
## polars

```{r}
#| label: df-string-sum-polars
pl$DataFrame(iris)$select(pl$sum("Petal.Length"))
```

## R base

```{r}
#| label: df-string-sum-rbase
sum(iris$Petal.Length)
```

## dplyr

```{r}
#| label: df-string-sum-dplyr
iris %>% summarise(sum(Petal.Length))
```

## data.table

```{r}
#| label: df-string-sum-datatable
sum(iris_dt[, Petal.Length])
```
:::


- From a `DataFrame` and a column as **an expression**:

```{r}
#| label: df-exp-sum-polars
pl$DataFrame(iris)$select(pl$sum(pl$col("Petal.Width")))
```

- From a `DataFrame` and a column as a list and **sum horizontally**:

In this case, use `with_column()` method.

::: {.panel-tabset}
## polars

```{r}
#| label: df-horizontally-sum-polars
df_pl <- pl$DataFrame(col1 = c(10L,20L), col2= c(30L,40L), col3 = c(40L,50L))
df_pl$with_column(pl$sum(list("col1", "col3")))
```

## R base

```{r}
#| label: df-horizontally-rbase
df <- data.frame(col1 = c(10L,20L), col2= c(30L,40L), col3 = c(40L,50L))
mysum <- rowSums(df[, c("col1", "col3")])
cbind(df,mysum)
```

## dplyr

```{r}
#| label: df-horizontally-sum-dplyr
df %>%
  rowwise() %>% 
  mutate(mysum = sum(col1,col3))
```

## data.table

```{r}
#| label: df-horizontally-sum-datatable
df_dt <- as.data.table(df)
df_dt[, somme := rowSums(.SD), .SDcols = c("col1", "col3")]
print(df_dt)
```
:::

- From a `DataFrame` and **sum horizontally** all columns:

In this case, use `list(*)`.

```{r}
#| label: df-list-sum-polars
df_pl$with_column(pl$sum(list("*")))
```

With iris `Dataframe`, you must first select numerical variables:

```{r}
#| label: df-list-sum-iris-polars
pl$DataFrame(iris)$
  select(
  pl$col(c("Petal.Length","Petal.Width")))$
  with_column(pl$sum(list("*")))
```
