## Strings methods

In polars, a lot of **strings methods** are useful.
To use them, simply prefix them with `str`.

### Get substrings

The `str$slice()` method can be used to create substrings of the string values 
of a **Utf8 Series**.  

`str$slice()` does not work like R base's `substr()` function for finding the substring of interest:
- `substr()` takes two arguments: the first and last elements;
- `str$slice()` takes two arguments: the first element and the extraction length.

::: {.callout-important}
With Polars, numeric default is 0! Thus the equivalent to `str$slice(0,3)` with Polars will be `substr(1,3)`.
:::

Two further comments:  

- If the second argument `length` is not specified, the sub-character string of interest will default to the end of the character string. For example in a `DataFrame` if `mycol` is a string column of length 4, `pl.col("mycol").str.slice(1)` is equivalent to `substr(mycol,2,4)` in dplyr.

- The first argument accepts negative values, which means that sub-strings can be considered starting from the end. For example in a `DataFrame` if `mycol` is a string column of length 4, `pl.col("mycol").str.slice(-2)` is equivalent to `substr(mycol,3,4)` in dplyr.

Let's see an example:

```{r}
#| label: creation-mydf-strings-example

mydf <- data.frame(
  col1 = 1:4,
  col2 = c("One_X","One_Y","Two_X","Two_Y")
)
```

::: {.panel-tabset}
## polars

```{r}
#| label: string-slice-polars
pl$DataFrame(mydf)$with_columns(
  pl$col("col2")$str$slice(0,length=3)$alias("level"),
  pl$col("col2")$str$slice(-1)$alias("x_y")
)
```

## R base

```{r}
#| label: string-slice-rbase
mydf$level <- substr(mydf$col2, 1, 3)
mydf$x_y <- substr(mydf$col2, nchar(mydf$col2), nchar(mydf$col2))
mydf
```

## dplyr

```{r}
#| label: string-slice-dplyr
mydf |>
  mutate(level = substr(col2, 1, 3),
         x_y = substr(col2, nchar(col2), nchar(col2)))
```

## data.table

```{r}
#| label: string-slice-datatable
mydt <- as.data.table(mydf) 
mydt[, c("level", "x_y") := .(substr(col2, 1, 3), substr(col2, nchar(col2), nchar(col2)))]
```
:::
