## Data Structures

> The core base data structures provided by Polars are Series and DataFrames. 

### Series and vectors

::: {.callout-important}
`Series` are a **1-dimensional** data structure. Within a series all elements have the same Data Type. 
:::

In Polars objects, `Series` object are like R `vectors`.  
To create a Polars Series from scratch:

::: {.panel-tabset}
## polars

```{r}
#| label: series-polars
mynumbers_serie <- pl$Series(1:3)
myletters_serie <- pl$Series(c("a","b","c"))
```

## R base

```{r}
#| label: vectors-rbase
mynumbers_vector <- 1:3
myletters_vector <- c("a","b","c")
```

:::

### DataFrame and data.frame

::: {.callout-note}
A `DataFrame` is a **2-dimensional** data structure that is backed by a `Series`, and it can be seen as an abstraction of a collection (e.g. list) of Series. 
:::

In polars objects, `DataFrame` object are like R `data.frame` and close to a `tibble` and a `data.table` object. `DataFrame` has some attributes and you can see [here](#initial-informations-on-a-dataframe) to know how you can use it.  

**To create a Polars DataFrame from scratch:**

::: {.panel-tabset}
## polars

```{r}
#| label: dataframe-polars
# Creation of a DataFrame object with Series
mydf <- pl$DataFrame(
  col1 = mynumbers_serie,
  col2 = myletters_serie
)
# Creation of a DataFrame object with Series and vectors
pl$DataFrame(
  col1 = mynumbers_serie,
  col2 = myletters_vector
)
```

## R base

```{r}
#| label: dataframe-rbase
data.frame(
  col1 = mynumbers_vector,
  col2 = myletters_vector
)
```

## dplyr

```{r}
#| label: tibble-dplyr
tibble(
  col1 = mynumbers_vector,
  col2 = myletters_vector
)
```

## data.table

```{r}
#| label: tibble-datatable
data.table(
  col1 = mynumbers_vector,
  col2 = myletters_vector
)
```

:::