## First glimpse

From [the official documentation](https://pola-rs.github.io/r-polars/vignettes/polars.html#series-and-dataframes):

> In polars, objects of class Series are analogous to R vectors. Objects of class DataFrame are analogous to R data frames. Notice that to avoid collision with classes provided by other packages, the class name of all objects created by polars starts with “RPolars”. For example, a polars DataFrame has the class “RPolarsDataFrame”.

> To create Polars Series and DataFrames objects, we load the library and use constructor functions with the pl$ prefix. This prefix is very important, as most of the polars functions are made available via pl$:

### Convert an existing R data.frame to a polars DataFrame

First example to convert the most famous R data frame ([iris](https://archive.ics.uci.edu/ml/datasets/iris)) to a Polars DataFrame.

To convert existing R data.frame to polars DataFrame, you can use `as_polars_df()` function :

```{r}
#| label: iris-conversion
library(polars)
iris_polars <- as_polars_df(iris)
iris_polars
```

### Count the number of lines

For example, **to count the number of lines** of the iris data frame :  

::: {.panel-tabset}
## polars

```{r}
#| label: height-function
# With pl$ prefix
as_polars_df(iris)$height
# Using iris_polars
iris_polars$height
```

## R base

```{r}
#| label: nrow-function
nrow(iris)
```

:::

### Extract data from a DataFrame

**To select the first 5 iris rows and the Petal.Length and Species columns**, syntax is identical between Polars and R base:

::: {.panel-tabset}
## polars

```{r}
#| label: first-extract-polars
iris_polars[1:5, c("Petal.Length", "Species")]
```

## tidypolars

```{r}
#| label: first-extract-tidypolars
iris_polars |>
  slice_head(n = 5) |>
  select(Petal.Length,Species)
```

## R base

```{r}
#| label: first-extract-rbase
iris[1:5, c("Petal.Length", "Species")]
```

## dplyr

```{r}
#| label: first-extract-dplyr
iris |> 
  slice_head(n = 5) |> 
  select(Petal.Length,Species)
```

## data.table

```{r}
#| label: first-extract-datatable
iris_dt[1:5, .(Petal.Length, Species)]
```
:::