## First glimpse

Polars' main functions are stored in the "pl" namespace and can be accessed using the "pl$" prefix to prevent conflicts with other packages and base R function names. For more, see [here](https://rpolars.github.io/reference/pl_pl/#examples).

### Convert a R data.frame to a polars DataFrame

First example to convert the most famous R data frame ([iris](https://archive.ics.uci.edu/ml/datasets/iris)) to a Polars DataFrame:

```{r}
#| label: iris-conversion
iris_polars <- pl$DataFrame(iris)
iris_polars
```

### Count the number of lines

For example, **to count the number of lines** of the iris data frame :  

::: {.panel-tabset}
## polars

```{r}
#| label: height-function
# With pl$ prefix
pl$DataFrame(iris)$height
# Using iris_polars
iris_polars$height
```

## R base

```{r}
#| label: nrow-function
nrow(iris)
```

:::

### Extract data from a DataFrame

**To select the first 5 iris rows and the Petal.Length and Species columns**, syntax is identical between Polars and R base:

::: {.panel-tabset}
## polars

```{r}
#| label: first-extract-polars
iris_polars[1:5, c("Petal.Length", "Species")]
```

## R base

```{r}
#| label: first-extract-rbase
iris[1:5, c("Petal.Length", "Species")]
```

## dplyr

```{r}
#| label: first-extract-dplyr
iris |> 
  slice(1:5) |> 
  select(Petal.Length,Species)
```

## data.table

```{r}
#| label: first-extract-datatable
iris_dt[1:5, .(Petal.Length, Species)]
```
:::