First of all we need to install all the packages and create an big random dataset needed for this book to work, so don't bother with the following code:

```{r}
#| label: necessary-settings
#| code-fold: true
#| message: false
#| warning: false
#| results: 'hide'

# Set CRAN mirror (required for non-interactive environments)
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Installation of packages for cookbook-rpolars
packages <- c('dplyr','data.table','tidyr','arrow','DBI','fakir','tictoc','duckdb','microbenchmark','readr','fs','ggplot2','pryr','dbplyr','forcats','collapse')
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages], dependencies = TRUE)
}

# Loading packages
invisible(lapply(packages, library, character.only = TRUE))

# Load tidypolars
library(tidypolars)

# Creation of iris_dt
iris_dt <- as.data.table(iris)
```

## Installation

Until the R polars package is uploaded to CRAN, [the polars package development team](https://github.com/pola-rs/r-polars/graphs/contributors) offers [several solutions](https://rpolars.github.io/#install) for installation.

The most practical one in my opinion at the moment is to use [R-universe](https://r-universe.dev/search/) and install like this:

```{.r}
install.packages("polars", repos = "https://community.r-multiverse.org")
```

To know the version of the `polars` package you have just installed and to have information on which features are enabled, you can use the `polars_info()` function.

```{r}
#| label: polars_info-function
library(polars)

polars_info()
```

If you want to install also the tidypolars package, you can do:

```{.r}
install.packages("tidypolars", repos = c("https://community.r-multiverse.org", 'https://cloud.r-project.org'))
