# Benchmarking

Below is the code used to create the data needed for the comparisons.  
Click to expand it! 👇

```{r}
#| label: load-packages-lazy-execution
#| code-fold: true
#| message: false
#| warning: false
#| results: 'hide'
library(polars)
library(arrow)
library(dplyr)
library(tictoc)

# Creation the "Datasets" folder
dir.create(normalizePath("Datasets"))

# Creation of large example R data.frame
DataMultiTypes <- data.frame(
  colInt = sample(1:2000000, 10000, replace = TRUE), 
  colNum = runif(2000000), 
  colString = sample(c("A", "B", "C"), 2000000, replace = TRUE), 
  colFactor = factor(sample(c("Low", "Medium", "High"), 2000000, replace = TRUE))
)

# Creation of large csv file
write.csv(x = DataMultiTypes,
          file = "Datasets/DataMultiTypes.csv")

# Creation of unique parquet file
arrow::write_parquet(x = DataMultiTypes,
                     sink = "Datasets/DataMultiTypes.parquet")
# Creation of partitionned parquet file
arrow::write_dataset(dataset = DataMultiTypes,
                     path = "Datasets/DataMultiTypes",
                     format = c("parquet"),
                     partitioning = c("colFactor"))
# Creation of duckdb file
con <- dbConnect(duckdb::duckdb(),
                 "Datasets/DataMultiTypes.duckdb")
duckdb::dbWriteTable(con,
                     "DataMultiTypes",
                     DataMultiTypes,
                     overwrite = TRUE)
dbDisconnect(con)
```

In this part of the book we will compare the performance of `polars` by comparing 
with other syntaxes, in particular `R base`, `dplyr`, `dbplyr`, `SQL` and `data.table`.

This section is structured according to the **type of file format** used for the comparison

## From a csv file

## From an unique parquet file

## From a partitionned parquet file

## From a duckdb file

PLAN :
1. A partir d'un fichier csv, comparer les performances d'une requête avec :
  i. R Base
  ii. dplyr
  iii. data.table
  iv. polars
2. A partir d'un fichier parquet unique, comparer : 
  i. dplyr
  ii. polars
3. A partir d'un fichier parquet partitionné, comparer : 
  i. dplyr
  ii. polars
4. La même comparaison à partir d'un fichier duckdb
  i. requête sql
  ii. dbplyr
  iii. polars


## Polars to DuckDB

DuckDB can natively query Polars DataFrames by referring to the name of Polars DataFrames as they exist in the current scope.


```{r}
con <- dbConnect(duckdb::duckdb(),
                 "Datasets/data.duckdb")
# Fonctionne
duckdb::dbWriteTable(con,
                     "iris_table",
                     iris,
                     overwrite = TRUE)
# Fonctionne
duckdb::dbWriteTable(con,
                     "mtcars_table",
                     mtcars,
                     overwrite = TRUE)
# Ne fonctionne pas
duckdb::dbWriteTable(con,
                     "iris_pl_table",
                     pl$DataFrame("iris"),
                     overwrite = TRUE)
dbListTables(con)
dbGetQuery(con, "SELECT * FROM iris_table WHERE SPECIES == 'setosa'")
dbExecute(con, "CREATE TABLE items(item VARCHAR, value DECIMAL(10,2), count INTEGER)")
dbExecute(con, "INSERT INTO items VALUES ('jeans', 20.0, 1), ('hammer', 42.2, 2)")
res <- dbGetQuery(con, "SELECT * FROM items")
print(res)
```


